---
title: "Wemix"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warnings = FALSE,
                      message = FALSE,
                      comment = "#>",
                      #results = "hide",
                      digits = 4,
                      error = FALSE)

## clean the R environment
graphics.off()
rm(list = ls())
freshr::freshr()

## load packages
library(here, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(gtsummary, quietly = TRUE)
library(flextable, quietly = TRUE)

library(lme4)
library(Matrix)
library(WeMix)

## check the directory for the file
# here::dr_here()
here::set_here()

## the figure or results should be saved 
# paste0("foldername/Sfilename_workingresult_", 
#      Sys.Date(), ".filetype")
```

```{css, echo=FALSE}
.scroll-100 {
  max-height: 300px;
  max-width: 1000px;
  overflow-y: auto;
  background-color: inherit;
}
```

::: callout-important
Matrix \< 1.6-2 and Matrix \>= 1.6-2 are binary incompatible. When you change between them, you must re-install from sources packages that link Matrix and therefore depend on the Matrix.

If not, then you need to install compatible binaries of Matrix and lme4 from a repository. Not all repositories provide compatible binaries, but CRAN does, so the following should work in most cases.

Matrix \< 1.6-2 and Matrix \>= 1.6-2 are binary incompatible. When you change between them, you must re-install from sources packages that link Matrix and therefore depend on the Matrix ABI:

> tools::package_dependencies("Matrix", which = "LinkingTo", reverse = TRUE)\[\[1L\]\] \[1\] "ahMLE" "bayesWatch" "cplm"\
> \[4\] "GeneralizedWendland" "hibayes" "irlba"\
> \[7\] "lme4" "mcmcsae" "OpenMx"\
> \[10\] "PRIMME" "robustlmm" "spGARCH"\
> \[13\] "TMB" "bcSeq"\
> If you have the required tools \[1\], then you can re-install lme4 from sources:

Note that binaries in all repositories will be rebuilt automatically once lme4 \> 1.1-35.1 is released. At that point, a simple update.packages() call will get things working again, even without explicit setting of repos.
:::

```{r}
# install.packages("WeMix")
# 
# oo <- options(repos = "https://cran.r-project.org/")
# install.packages("Matrix")
# install.packages("lme4")
# options(oo)

```

Standard errors of random effect variances are robust

```{r}
data(sleepstudy)
ss1 <- sleepstudy

# Create weights
ss1$W1 <- ifelse(ss1$Subject %in% c(308, 309, 310), 2, 1)
ss1$W2 <- 1

View(sleepstudy)
ss1

# Run random-intercept 2-level model 
two_level <- mix(Reaction ~ Days + (1|Subject), 
                 data = ss1, weights = c("W1", "W2"))

#Run random-intercept 2-level model with group-mean centering
grp_centered <- mix(Reaction ~ Days + (1|Subject), 
                    data = ss1,
                    weights = c("W1", "W2"),
                    verbose = TRUE,
                    center_group = list("Subject" = ~Days))
```

```{r}
summary(grp_centered)
```

```{r}
#Run three level model with random slope and intercept. 
#add group variables for 3 level model 
ss1$Group <- 3
ss1$Group <- ifelse(as.numeric(ss1$Subject) %% 10 < 7, 2, ss1$Group)
ss1$Group <- ifelse(as.numeric(ss1$Subject) %% 10 < 4, 1, ss1$Group)
# level-3 weights
ss1$W3 <- ifelse(ss1$Group == 2, 2, 1)
```

```{r, class.output="scroll-100"}
ss1
```

```{r}
three_level <- mix(Reaction ~ Days + (1 | Subject) + (1 + Days | Group),
                   data = ss1,
                   weights = c("W1", "W2", "W3"))
```

```{r, class.output="scroll-100"}
summary(three_level)
```
